---
title: "Lab 11"
author: "Andrew Douglass"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tasks

## Task 1

```{r}
getwd()
```


## Task 2

```{r}
#Given data
d = c(5.0581, 4.9707, 5.0893, 4.9334, 4.9777, 5.0285, 4.8555, 4.9565, 
       4.9769, 4.9722, 4.999, 4.9925, 4.9686, 5.0662, 4.9239, 4.9781, 
       5.0485, 5.0014, 4.9957, 5.0195, 5.0118, 4.9928, 5.0361, 5.0185, 
       4.9879)

#shared information for all confidence intervals
dmean = mean(d)
dSD = sd(d)
n = length(d)

#95%
tvalue95 = qt(0.975, n - 1)
ciValue95 = c(dmean - tvalue95 * dSD / sqrt(n), dmean + tvalue95 * dSD / sqrt(n))
cat("95% CI:", ciValue95, "\n")

#90%
tvalue90 = qt(0.95, n - 1)
ciValue90 = c(dmean - tvalue90 * dSD / sqrt(n), dmean + tvalue90 * dSD / sqrt(n))
cat("90% CI:", ciValue90, "\n")

#80%
tvalue80 = qt(0.90, n - 1)
ciValue80 = c(dmean - tvalue80 * dSD / sqrt(n), dmean + tvalue80 * dSD / sqrt(n))
cat("80% CI:", ciValue80, "\n")

#50%
tvalue50 = qt(0.75, n - 1)
ciValue50 = c(dmean - tvalue50 * dSD / sqrt(n), dmean + tvalue50 * dSD / sqrt(n))
cat("50% CI:", ciValue50, "\n")

```

```{r}
d = c(5.0581, 4.9707, 5.0893, 4.9334, 4.9777, 5.0285, 4.8555, 4.9565, 
       4.9769, 4.9722, 4.999, 4.9925, 4.9686, 5.0662, 4.9239, 4.9781, 
       5.0485, 5.0014, 4.9957, 5.0195, 5.0118, 4.9928, 5.0361, 5.0185, 
       4.9879)

obj = t.test(d, conf.level = 0.80)
ci80 = obj$conf.int
cat("80% ci for the mean using t.test():", ci80, "\n")

#Using bootstrapping to calculate the confidence intervals
bootSamples = 10000

#Storing the variances
bootstrapVariance = numeric(bootSamples)

set.seed(123)

#Creating bootstrap samples and calculating the variance
for (i in 1:bootSamples) {
  bootstrap_sample = sample(d, replace = TRUE)
  bootstrapVariance[i] = var(bootstrap_sample)
}

# Calculate confidence intervals by finding percentiles
civar95 = quantile(bootstrapVariance, probs = c(0.025, 0.975))
civar90 = quantile(bootstrapVariance, probs = c(0.05, 0.95))
civar80 = quantile(bootstrapVariance, probs = c(0.10, 0.90))
civar50 = quantile(bootstrapVariance, probs = c(0.25, 0.75))

cat("95% CI:", civar95, "\n")
cat("90% CI:", civar90, "\n")
cat("80% CI:", civar80, "\n")
cat("50% CI:", civar50, "\n")

```

## Task 3

```{r}
#given
blue = c(21.65, 17.48, 20.1, 21.57, 14.82, 19.17, 21.08, 18.23, 22.93, 15.66, 20.89, 21.66, 18.5, 20.59, 18.63, 18.91, 19.53, 17.7, 16.5, 19.03)
snapper = c(31.65, 27.48, 30.1, 31.57, 24.82, 29.17, 31.08, 28.23, 32.93, 25.66, 30.89, 31.66, 28.5, 30.59, 28.63)

#Getting sample lengths, mean, variance, pooled variance, t-value, then calculating the confidence interval
n1 = length(blue)
n2 = length(snapper)

mblue = sum(blue) / n1
msnapper = sum(snapper) / n2

varblue = sum((blue - mblue)^2) / (n1 - 1)
varsnapper = sum((snapper - msnapper)^2) / (n2 - 1)

pooled = ((n1 - 1) * varblue + (n2 - 1) * varsnapper) / (n1 + n2 - 2)

alpha = 0.05
talpha = qt(1 - alpha / 2, df = n1 + n2 - 2)

#CI calculation
error = talpha * sqrt(pooled * (1 / n1 + 1 / n2))
ciLower = (mblue - msnapper) - error
ciUpper = (mblue - msnapper) + error

ci = c(ciLower, ciUpper)
cat("95% CI:", ci, "\n")

ciLowerSnapper_blue = (msnapper - mblue) - error
ciUpperSnapper_blue = (msnapper - mblue) + error

ciSnapper_blue = c(ciLowerSnapper_blue, ciUpperSnapper_blue)
cat("95% CI for snapper - blue:", ciSnapper_blue, "\n")

```
The first part of this output is for the difference in means of blue minus snapper. The values -11.84 and -8.76 indicate that since both values are negative then the mean length of blue is less than mean of snapper so the snapper have a higher chance of being longer.

The second part is where the mean difference of snapper - blue is performed. The value is 8.76 and 11.84 which indicates that the snapper have a higher chance of being longer than the blue. As its basically the opposite of the first part.

```{r}
blue = c(21.65, 17.48, 20.1, 21.57, 14.82, 19.17, 21.08, 18.23, 22.93, 
         15.66, 20.89, 21.66, 18.5, 20.59, 18.63, 18.91, 19.53, 17.7, 
         16.5, 19.03)
snapper = c(31.65, 27.48, 30.1, 31.57, 24.82, 29.17, 31.08, 28.23, 32.93, 
            25.66, 30.89, 31.66, 28.5, 30.59, 28.63)

ci95 = t.test(snapper, blue, var.equal=TRUE, conf.level=0.95)$conf.int
ci85 = t.test(snapper, blue, var.equal=TRUE, conf.level=0.85)$conf.int
ci75 = t.test(snapper, blue, var.equal=TRUE, conf.level=0.75)$conf.int
ci25 = t.test(snapper, blue, var.equal=TRUE, conf.level=0.25)$conf.int


cat("95% CI:", ci95, "\n")
cat("85% CI:", ci85, "\n")
cat("75% CI:", ci75, "\n")
cat("25% CI:", ci25, "\n")

```

As the confidence level decreases the interval would become smaller. This is because its making a more precise interval which is smaller and the only way to make it smaller would be by reducing the confidence in the result. A higher confidence can be made if a very wide interval is used because there is a much higher chance of the sample being in that interval which gives the higher confidence. 

## Task 4

```{r}
#given
Exam1 = c(40.98, 59.36, 46.69, 41.8, 61.63, 65.31, 62.96, 60.21, 56.89, 
           78.41, 53.44, 75.2, 60.54, 52.43, 41.41, 70.79, 73.55, 55.65, 
           61.43, 63.84, 58.07, 53.79, 54.45, 67.18, 44.46)

Exam2 = c(50.22, 66.19, 58.75, 51.88, 66.61, 70.86, 74.25, 70.23, 69.55, 
           87.18, 63.62, 81.7, 70.5, 66.02, 51.35, 80.92, 85.65, 65.44, 
           74.37, 75.28, 67.86, 59.92, 64.42, 73.57, 57.15)

#Exam differences, lengths, mean, t-value, and standard deviation for the confidence interval calculation
examDifference = Exam1 - Exam2

n = length(examDifference)

meanDifference = sum(examDifference) / n

SdDifference = sqrt(sum((examDifference - meanDifference)^2) / (n - 1))

alpha = 0.05
t_alpha = qt(1 - alpha / 2, df = n - 1)

error = t_alpha * (SdDifference / sqrt(n))
ciLower = meanDifference - error
ciUpper = meanDifference + error
ci_95 = c(ciLower, ciUpper)

cat("95% CI for mean difference:", ci_95, "\n")

```
Based on this CI interval of -10.73 and -8.71, since the difference is by Exam1 - Exam2 then this result indicates that the exam 1 scores are lower than exam 2's. Looking at the scores in Exam1 and Exam2 this makes sense as exam 2 has more higher scores such as in the 80s and 70s while exam 1 has more lower values in the 40s. On average, exam 1 should have between 8.7 and 10.7 lower points than from exam 2. 

```{r}
#t.test function 
ci_90 = t.test(examDifference, conf.level = 0.90)$conf.int
ci_80 = t.test(examDifference, conf.level = 0.80)$conf.int
ci_70 = t.test(examDifference, conf.level = 0.70)$conf.int
ci_60 = t.test(examDifference, conf.level = 0.60)$conf.int
ci_10 = t.test(examDifference, conf.level = 0.10)$conf.int

cat("90% CI:", ci_90, "\n")
cat("80% CI:", ci_80, "\n")
cat("70% CI:", ci_70, "\n")
cat("60% CI:", ci_60, "\n")
cat("10% CI:", ci_10, "\n")

```
As the confidence level decreases, the range of the interval becomes smaller which is in line with the comparison made in task 3.

## Task 5

```{r}
NZBIRDS = read.csv("NZBIRDS.csv")

total_species = table(NZBIRDS$Extinct)
#using with to check the counts of flightless and extinct birds
flightless_counts = with(NZBIRDS, table(Extinct, Flight))

extinct_total = total_species["Yes"]
non_extinct_total = total_species["No"]

extinct_flightless = flightless_counts["Yes", "No"]
non_extinct_flightless = flightless_counts["No", "No"]

#printing values for comparison
cat("Extinct Total Species:", extinct_total, "\n")
cat("Extinct Flightless Species:", extinct_flightless, "\n")
cat("Non-Extinct Total Species:", non_extinct_total, "\n")
cat("Non-Extinct Flightless Species:", non_extinct_flightless, "\n")

#Values do match the table

p_extinct_flightless = extinct_flightless / extinct_total
p_non_extinct_flightless = non_extinct_flightless / non_extinct_total

sq = sqrt((p_extinct_flightless * (1 - p_extinct_flightless) / extinct_total) +
           (p_non_extinct_flightless * (1 - p_non_extinct_flightless) / non_extinct_total))

pd = p_extinct_flightless - p_non_extinct_flightless

z = 1.96

ci = c(pd - sq * z, pd + sq * z)

cat("95% CI:", ci, "\n")

```

## Task 6

```{r}

set.seed(35);sam1 = rnorm(25, mean = 10, sd = 5)
set.seed(45);sam2 = rnorm(34, mean = 40, sd = 8)


a = 1 - 0.05 / 2
ci_95 = c()
ci_95[1] = var(sam1) / (var(sam2) * qf(a, df1 = 24, df2 = 33))
ci_95[2] = var(sam1) / var(sam2) * qf(a, df1 = 33, df2 = 24)
cat("95%:", ci_95, "\n")

a = 1 - 0.20 / 2
ci_80 = c()
ci_80[1] = var(sam1) / (var(sam2) * qf(a, df1 = 24, df2 = 33))
ci_80[2] = var(sam1) / var(sam2) * qf(a, df1 = 33, df2 = 24)
cat("80%:", ci_80, "\n")

a = 1 - 0.30 / 2
ci_70 = c()
ci_70[1] = var(sam1) / (var(sam2) * qf(a, df1 = 24, df2 = 33))
ci_70[2] = var(sam1) / var(sam2) * qf(a, df1 = 33, df2 = 24)
cat("70%:", ci_70, "\n")

a = 1 - 0.40 / 2
ci_60 = c()
ci_60[1] = var(sam1) / (var(sam2) * qf(a, df1 = 24, df2 = 33))
ci_60[2] = var(sam1) / var(sam2) * qf(a, df1 = 33, df2 = 24)
cat("60%:", ci_60, "\n")

a = 1 - 0.50 / 2
ci_50 = c()
ci_50[1] = var(sam1) / (var(sam2) * qf(a, df1 = 24, df2 = 33))
ci_50[2] = var(sam1) / var(sam2) * qf(a, df1 = 33, df2 = 24)
cat("50%:", ci_50, "\n")

```

## Task 7

```{r}
library(FALL24DOUG0060)
set.seed(23)
x = rnorm(30, mean = 10, sd = 12)
FALL24DOUG0060::myci(x)

```



