---
title: "Lab 10"
author: "Andrew Douglass"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tasks

## Task 1

```{r}
getwd()
```


## Task 2

```{r}
mymaxlik=function(lfun,x,param,...){
# how many param values are there?
np=length(param)
# outer -- notice the order, x then param
# this produces a matrix -- try outer(1:4,5:10,function(x,y) paste(x,y,sep=" "))   to understand
z=outer(x,param,lfun) # A
# z is a matrix where each x,param is replaced with the function evaluated at those values
y=apply(z,2,sum)

# y is a vector made up of the column sums
# Each y is the log lik for a new parameter value
plot(param,y,col="Blue",type="l",lwd=2,...)
# which gives the index for the value of y == max.
# there could be a max between two values of the parameter, therefore 2 indices
# the first max will take the larger indice
i=max(which(y==max(y))) # B
abline(v=param[i],lwd=2,col="Red")

# plots a nice point where the max lik is
points(param[i],y[i],pch=19,cex=1.5,col="Black")
axis(3,param[i],round(param[i],2))
#check slopes. If it is a max the slope shoud change sign from + to 
# We should get three + and two -vs
ifelse(i-3>=1 & i+2<=np, slope<-(y[(i-2):(i+2)]-y[(i-3):(i+1)])/(param[(i-2):(i+2)]-param[(i-3):(i+1)]),slope<-"NA")
return(list(i=i,parami=param[i],yi=y[i],slope=slope))
}
```

Line A creates a matrix by using the function lfun for all the x values and parameter combinations. The Outer function is what creates the matrix and computes the value of lfun for all combinations of x and param.

Line B finds the index of the maximum likelihood for y. For where the maximum value of y is found, the which function returns all indicies where the value is equal to the max in y. The largest index should be chosen because of the use of max.

Formula for likelihood:

Starts as:
$$
L(p) = P(Y_1 = y_1) \cdot P(Y_2 = y_2) \ldots \cdot P(Y_k = y_k)
$$
For this problem it will then become the following based on the question 7.8 hint.
$$
L(p) = p^{\sum_{i=1}^{8} y_i} (1 - p)^{8 \cdot 20 - \sum_{i=1}^{8} y_i}
$$

R Code for the function and finding the maximum likelihood estimate using y = 3,3,4,3,4,5,5,4:
```{r}
y <- c(3, 3, 4, 3, 4, 5, 5, 4)

p <- 0.2

Y <- sum(y)

n <- length(y) * 20

L <- p^Y * (1 - p)^(n - Y)

L

```

## Task 3

```{r}
logpoiss=function(x,param) log(dpois(x,lambda=param)) 
mymaxlik=function(lfun,x,param,...){
# how many param values are there?
np=length(param)
# outer -- notice the order, x then param
# this produces a matrix -- try outer(1:4,5:10,function(x,y) paste(x,y,sep=" "))   to understand
z=outer(x,param,lfun) # A
# z is a matrix where each x,param is replaced with the function evaluated at those values
y=apply(z,2,sum)

# y is a vector made up of the column sums
# Each y is the log lik for a new parameter value
plot(param,y,col="Blue",type="l",lwd=2,...)
# which gives the index for the value of y == max.
# there could be a max between two values of the parameter, therefore 2 indices
# the first max will take the larger indice
i=max(which(y==max(y))) # B
abline(v=param[i],lwd=2,col="Red")

# plots a nice point where the max lik is
points(param[i],y[i],pch=19,cex=1.5,col="Black")
axis(3,param[i],round(param[i],2))
#check slopes. If it is a max the slope shoud change sign from + to 
# We should get three + and two -vs
ifelse(i-3>=1 & i+2<=np, slope<-(y[(i-2):(i+2)]-y[(i-3):(i+1)])/(param[(i-2):(i+2)]-param[(i-3):(i+1)]),slope<-"NA")
return(list(i=i,parami=param[i],yi=y[i],slope=slope))
}

y <- c(4, 6, 7, 6, 5)

lambda_values <- seq(0, 10, length.out = 1000)

mymaxlik(lfun = logpoiss, x = y, param = lambda_values, xlab = expression(lambda), main = "Poisson")

```

Based on:
$$
P(y) = \frac{e^{-\lambda} \lambda^y}{y!}
$$
The following equation can be simplified to:
$$
\ln L(\lambda) = Y \ln(\lambda) - 5\lambda
$$

```{r}
myNRML=function(x0,delta=0.001,llik,xrange,parameter="param"){
f=function(x) (llik(x+delta)-llik(x))/delta
fdash=function(x) (f(x+delta)-f(x))/delta
d=1000
i=0
x=c()
y=c()
x[1]=x0
y[1]=f(x[1])
while(d > delta & i<100){
i=i+1
x[i+1]=x[i]-f(x[i])/fdash(x[i])
y[i+1]=f(x[i+1])
d=abs(y[i+1])
}
layout(matrix(1:2,nr=1,nc=2,byrow=TRUE),width=c(1,2))
curve(llik(x), xlim=xrange,xlab=parameter,ylab="log Lik",main="Log Lik")
curve(f(x),xlim=xrange,xaxt="n", xlab=parameter,ylab="derivative",main=  "Newton-Raphson Algorithm \n on the derivative")
points(x,y,col="Red",pch=19,cex=1.5)
axis(1,x,round(x,2),las=2)
abline(h=0,col="Red")

segments(x[1:(i-1)],y[1:(i-1)],x[2:i],rep(0,i-1),col="Blue",lwd=2)
segments(x[2:i],rep(0,i-1),x[2:i],y[2:i],lwd=0.5,col="Green")

list(x=x,y=y)
}

y <- c(4, 6, 7, 6, 5)

llik <- function(lambda) {
  sum(y) * log(lambda) - length(y) * lambda
}

result <- myNRML(x0 = 1, delta = 0.000001, llik = llik, xrange = c(0.01, 20), parameter = "lambda")

lambda_mle <- result$x[length(result$x)]
cat("lambda:", lambda_mle, "\n")


```

## Task 4

```{r}
mymaxlikg=function(lfun="logbin2",theta) { # default log lik is a combination bin
nth=length(theta)  # nu. of valuse used in theta
thmat=matrix(theta,nr=nth,nc=1,byrow=TRUE) # Matrix of theta
z=apply(thmat,1,lfun) # z holds the log lik values
zmax=max(which(z==max(z)))  # finding the INDEX of the max lik
plot(theta,exp(z),type="l") # plot of lik
abline(v=theta[zmax],col="Blue")   #  verical line through max
axis(3,theta[zmax],round(theta[zmax],4))  # one tick on the third axis 
theta[zmax]   # theta corresponding to max lik
}

logbin2 <- function(theta) {
  log(dbinom(2, size = 6, prob = theta)) + log(dbinom(4, size = 10, prob = theta))
}

mymaxlikg(theta = seq(0, 1, length.out = 10000))
#Main adjustment in logbin2 for new data.
```

## Task 5

Based on the given equation:
$$
p(y_1, y_2 \mid \theta_1, \theta_2) = \text{bin}(y_1 \mid \theta_1) \cdot \text{poiss}(y_2 \mid \theta_2)
$$

The following equation can be made for the log likelihood:
$$
\ell(\theta_1, \theta_2) = \ln(p(y_1, y_2 \mid \theta_1, \theta_2))
$$
It takes the log of the side where p is as that side equals the binomial and Poisson side. It can then be extended out to form:
$$
\ell(\theta_1, \theta_2) = \ln \left( \binom{n}{y_1} \right) + y_1 \ln(\theta_1) + (n - y_1) \ln(1 - \theta_1) + y_2 \ln(\theta_2) - \theta_2 - \ln(y_2!)
$$
```{r}
maxlikg2=function(theta1,theta2,lfun="logbinpois",...){
n1=length(theta1)
n2=length(theta2)
z=outer(theta1,theta2,lfun)
contour(theta1,theta2,exp(z),...) # exp(z) gives the lik
maxl=max(exp(z))    # max lik
coord=which(exp(z)==maxl,arr.ind=TRUE)  # find the co-ords of the max
th1est=theta1[coord[1]] # mxlik estimate of theta1
th2est=theta2[coord[2]]
abline(v=th1est,h=th2est)
axis(3,th1est,round(th1est,2))
axis(4,th2est,round(th2est,2),las=1)
list(th1est=th1est,th2est=th2est)
}

logbinpois <- function(theta1, theta2) {
  log(dbinom(4, size = 20, prob = theta1)) +
  log(dpois(4, lambda = theta2))
}

theta1_values <- seq(0, 1, length.out = 1000)
theta2_values <- seq(0, 10, length.out = 1000)

result <- maxlikg2(theta1 = theta1_values, theta2 = theta2_values, lfun = "logbinpois", nlevels = 20)

cat("Theta1:", result$th1est, "\n")
cat("Theta2:", result$th2est, "\n")


```

## Task 6

```{r}
mymlnorm=function(x,mu,sig,...){  #x sample vector
nmu=length(mu) # number of values in mu
nsig=length(sig)
n=length(x) # sample size
zz=c()    ## initialize a new vector
lfun=function(x,m,p) log(dnorm(x,mean=m,sd=p))   # log lik for normal
for(j in 1:nsig){
z=outer(x,mu,lfun,p=sig[j]) # z a matrix 
# col 1 of z contains lfun evaluated at each x with first value of mu, 
# col2 each x with 2nd value of m 
# all with sig=sig[j]
y=apply(z,2,sum)
# y is a vector filled with log lik values, 
# each with a difft mu and all with the same sig[j]
zz=cbind(zz,y)
## zz is the matrix with each column containing log L values, rows difft mu, cols difft sigmas 
}
maxl=max(exp(zz))
coord=which(exp(zz)==maxl,arr.ind=TRUE)
maxlsig=apply(zz,1,max)
contour(mu,sig,exp(zz),las=3,xlab=expression(mu),ylab=expression(sigma),axes=TRUE,
main=expression(paste("L(",mu,",",sigma,")",sep="")), labcex = .8,...) #Made internal graph numbers bigger
mlx=round(mean(x),2)  # theoretical
mly=round(sqrt((n-1)/n)*sd(x),2)
#axis(1,at=c(0:20,mlx),labels=sort(c(0:20,mlx)))
#axis(2,at=c(0:20,mly),labels=TRUE)
abline(v=mean(x),lwd=2,col="Green")
abline(h=sqrt((n-1)/n)*sd(x),lwd=2,col="Red")

# Now find the estimates from the co-ords
muest=mu[coord[1]]
sigest=sig[coord[2]]

abline(v=muest, h=sigest)
return(list(x = x, coord = coord, maxl = maxl, muest = muest, sigest = sigest))
}

y <- c(10, 12, 13, 15, 12, 11, 10)

mu_values <- seq(5, 20, length = 1000)
sigma_values <- seq(0.1, 10, length = 1000)

result <- mymlnorm(x = y, mu = mu_values, sig = sigma_values)

cat("mu:", result$muest, "\n")
cat("sigma:", result$sigest, "\n") #Adjusted the return statement to output these values

```

## Task 7

```{r}
mymlbeta=function(x,alpha,beta,...){  #x sample vector
na=length(alpha) # number of values in alpha
nb=length(beta)
n=length(x) # sample size
zz=c()    ## initialize a new vector
lfun=function(x,a,b) log(dbeta(x,shape1=a,shape2=b))   # log lik for beta
for(j in 1:nb){
z=outer(x,alpha,lfun,b=beta[j]) # z a matrix 
# col 1 of z contains lfun evaluated at each x with first value of alpha, 
# col2 each x with 2nd value of a 
# all with b=beta[j]
y=apply(z,2,sum)
# y is a vector filled with log lik values, 
# each with a difft alpha and all with the same sig[j]
zz=cbind(zz,y)
## zz is the matrix with each column containing log L values, rows difft alpha, cols difft betas 
}
maxl=max(exp(zz))    # max lik
coord=which(exp(zz)==maxl,arr.ind=TRUE)  # find the co-ords of the max
aest=alpha[coord[1]] # mxlik estimate of alpha
best=beta[coord[2]]
contour(alpha,beta,exp(zz),las=3,xlab=expression(alpha),ylab=expression(beta),axes=TRUE,
main=expression(paste("L(",alpha,",",beta,")",sep="")),...)

abline(v=aest, h=best)
points(aest,best,pch=19)
axis(4,best,round(best,2),col="Red")
axis(3,aest,round(aest,2),col="Red")
return(list(x=x,coord=coord,maxl=maxl,maxalpha=aest,maxbeta=best))
}

set.seed(123) #Recreate same results
sam <- rbeta(30, shape1 = 3, shape2 = 4) #Creating sample from given data

#parameters
givenShape1 <- 3
givenShape2 <- 4

#12 Plots on same section
par(mfrow = c(3, 4))

#Resampling 12 times
for (i in 1:12) {
  resample <- sample(sam, size = 30, replace = TRUE)
  
  #Using the mymlbeta function to find the maximum likelihood estimate
  result <- mymlbeta(
    x = resample,
    alpha = seq(1, 5, length = 100),
    beta = seq(2, 6, length = 100)
  )

  #Plotting contour plot and adding details
  abline(v = givenShape1, h = givenShape2, col = "red", lty = 2)
  points(result$maxalpha, result$maxbeta, pch = 19, col = "blue")
}


```


## Task 8

```{r}
library(FALL24DOUG0060)

sample_data <- rbeta(10, shape1 = 2, shape2 = 3)

result <- FALL24DOUG0060::mymlbeta(
  x = sample_data,
  alpha = seq(1, 4, length = 100),
  beta = seq(2, 5, length = 100)
)
```


