---
title: "Lab 12"
author: "Andrew Douglass"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tasks

## Task 1

```{r}
getwd()
```


## Task 2

```{r}
set.seed(55);x1=rnorm(30,mean=25,sd=5)
h0=t.test(x1,mu=22)
h0$p.value; h0$conf.int
```

The p-value is less than 0.05 so H0 is rejected in favor of H1.

```{r}
h02=t.test(x1,mu=23)
h02$p.value; h02$conf.int
```
The p-value is also less than 0.05 so it is rejected in favor of H1.

```{r}
h03=t.test(x1,mu=24)
h03$p.value; h03$conf.int
```
The p-value is greater than 0.05 and mu is in the confidence interval so H0 is not rejected.

```{r}
h04=t.test(x1,mu=25)
h04$p.value; h04$conf.int
```

The p-value is greater than 0.05 and mu is in the confidence interval so H0 is not rejected.

```{r}
h05=t.test(x1,mu=26)
h05$p.value; h05$conf.int
```
The p-value is greater than 0.05 and mu is in the confidence interval so H0 is not rejected.

```{r}
#Make boxplot of the data and plot the sample mean, and the 95% confidence interval
boxplot(x1, main="Sample x1")

t.test(x1,mu=28)
ci=t.test(x1,mu=23)$conf.int
ci
abline(h=c(ci,mean(x1)),col=c("Red","Red","Green"))
```

```{r}
tcalc = (mean(x1)-24)/(sd(x1)/sqrt(30))
tcalc

mypvalue=function(t0,xmax=4,n=20, alpha=0.05){
#calculate alpha/2
va=round(pt(-t0,df=n-1),4)
pv=2*va

# plot the t dist
curve(dt(x,df=n-1),xlim=c(-xmax,xmax),ylab="T Density",xlab=expression(t),
main=substitute(paste("P-value=", pv, " alpha=", alpha)))

# set up points on the polygon to the right
xcurve=seq(t0,xmax,length=1000)
ycurve=dt(xcurve,df=n-1)

# set up points to the left
xlcurve=seq(-t0,-xmax,length=1000)
ylcurve=dt(xcurve,df=n-1)

# Shade in the polygon defined by the line segments
polygon(c(t0,xcurve,xmax),c(0,ycurve,0),col="green")
polygon(c(-t0,xlcurve,-xmax),c(0,ylcurve,0),col="green")

# make quantiles
q=qt(1-alpha/2,n-1)
abline( v=c(q,-q),lwd=2) # plot the cut off t value 
axis(3,c(q,-q),c(expression(abs(t[alpha/2])),expression(-abs(t[alpha/2]))))


# Annotation
text(0.5*(t0+xmax),max(ycurve),substitute(paste(area, "=",va)))
text(-0.5*(t0+xmax),max(ycurve),expression(area))

return(list(q=q,pvalue=pv))
}

mypvalue(tcalc, n=30, alpha=0.05)

```

Between the vertical lines would be the accepting region for the null hypothesis. Outside of the regions would be for not accepting. The green areas correlate to the p-values. The area where these meet as in the green area between the two lines would be the main area for accepting or rejecting H0. Since tcalc is 1.33 and the p-value is 0.2, we would not reject H0.

```{r}
bootpval<-function(x,conf.level=0.95,iter=3000,mu0=0, test="two"){
n=length(x)
y=x-mean(x)+mu0  # transform the data so that it is centered at the NULL
rs.mat<-c()    #rs.mat will become a resample matrix -- now it is an empty vector
xrs.mat<-c()
for(i in 1:iter){ # for loop - the loop will go around iter times
rs.mat<-cbind(rs.mat,sample(y,n,replace=TRUE)) #sampling from y cbind -- column bind -- binds the vectors together by columns
xrs.mat<-cbind(xrs.mat,sample(x,n,replace=TRUE)) #sampling from x cbind -- column bind -- binds the vectors together by columns

}

tstat<-function(z){ # The value of t when the NULL is assumed true (xbar-muo)/z/sqrt(n)
sqrt(n)*(mean(z)-mu0)/sd(z)
}

tcalc=tstat(x) # t for the data collected
ytstat=apply(rs.mat,2,tstat) # tstat of resampled y's, ytstat is a vector and will have iter values in it
xstat=apply(xrs.mat,2,mean)  # mean of resampled x's
alpha=1-conf.level # calculating alpha
ci=quantile(xstat,c(alpha/2,1-alpha/2))# Nice way to form a confidence interval
pvalue=ifelse(test=="two",length(ytstat[ytstat>abs(tcalc) | ytstat < -abs(tcalc)])/iter,
ifelse(test=="upper",length(ytstat[ytstat>tcalc])/iter,
length(ytstat[ytstat<xstat])/iter))

h=hist(ytstat,plot=FALSE)
mid=h$mid
if(test=="two"){
ncoll=length(mid[mid<= -abs(tcalc)])
ncolr=length(mid[mid>=  abs(tcalc)])
col=c(rep("Green",ncoll),rep("Gray",length(mid)-ncoll-ncolr),rep("Green",ncolr))
}
if(test=="upper"){
ncolr=length(mid[mid>=  abs(tcalc)])
col=c(rep("Gray",length(mid)-ncolr),rep("Green",ncolr))
}

if(test=="lower"){
ncoll=length(mid[mid<=  -abs(tcalc)])
col=c(rep("Green",ncoll),rep("Gray",length(mid)-ncoll))
}
hist(ytstat,col=col,freq=FALSE,las=1,main="",xlab=expression(T[stat]))
#segments(ci[1],0,ci[2],0,lwd=2)
pround=round(pvalue,4)
title(substitute(paste(P[value],"=",pround)))
return(list(pvalue=pvalue,tcalc=tcalc,n=n,x=x,test=test,ci=ci))
}


boot=bootpval(x=x1,mu0=22,test="two")
boot23=bootpval(x=x1,mu0=23,test="two")
boot24=bootpval(x=x1,mu0=24,test="two")
boot25=bootpval(x=x1,mu0=25,test="two")
boot26=bootpval(x=x1,mu0=26,test="two")
```

When the mu = 24, the resulting p-value is closer to the mypvalue function result and has a similar graph. Comparing these results to the initial H0 comparisons, the p-values are similar to the t-test results. Therefore the rejections made earlier are true in that they should.

## Task 3

```{r}
set.seed(30);x=rnorm(15,mean=10,sd=7)   
set.seed(40); y=rnorm(20,mean=12,sd=4)

var.test(x,y)
```
At first look, the p-value is less than the significance level of 0.05 so we would reject H0. This means that var.equal is false and the variances are not equal. 1 is also not contained in the confidence interval.

```{r}
t.test(y,x,mu=0,var.equal=FALSE)

t.test(y,x,mu=2,var.equal=FALSE)
```
Both of the p-values are at an acceptable amount and the confidence interval is wide enough that 1 will be included in it. Therefore for both of these tests, the difference between their means of x and y does not have evidence to prove its not 0 or 2 respectively. 

## Task 4

```{r}
set.seed(30);x=rnorm(15,mean=10,sd=4)
set.seed(40);y=rnorm(20,mean=12,sd=4)

var.test(x,y)

```
Given the p-value and confidence interval, the var.equal would be TRUE.

```{r}
t.test(y, x, mu=0, var.equal=TRUE)
t.test(y, x, mu=2, var.equal=TRUE)
```
The p-value between these two t.tests with different mean values are very different. However, the confidence intervals are very similar as well as the mean of x and the mean of y.

## Task 5

```{r}
boot2pval<-function(x1,x2,conf.level=0.95,iter=3000,mudiff=0, test="two"){
n1=length(x1)
n2=length(x2)
y1=x1-mean(x1)+mean(c(x1,x2))  # transform the data so that it is centered at the NULL
y2=x2-mean(x2)+mean(c(x1,x2))
y1rs.mat<-c()    #rs.mat will be come a resample matrix -- now it is an empty vector
x1rs.mat<-c()
y2rs.mat<-c()
x2rs.mat<-c()
for(i in 1:iter){ # for loop - the loop will go around iter times
y1rs.mat<-cbind(y1rs.mat,sample(y1,n1,replace=TRUE)) #sampling from y cbind -- column bind -- binds the vectors together by columns
y2rs.mat<-cbind(y2rs.mat,sample(y2,n2,replace=TRUE))

}
x1rs.mat<-y1rs.mat+mean(x1)-mean(c(x1,x2))
x2rs.mat<-y2rs.mat+mean(x2)-mean(c(x1,x2))

xbar1=mean(x1)
xbar2=mean(x2)
sx1sq=var(x1)
sx2sq=var(x2)

tcalc=(xbar1-xbar2-mudiff)/sqrt(sx1sq/n1+sx2sq/n2)

sy1sq=apply(y1rs.mat,2,var)
sy2sq=apply(y2rs.mat,2,var) 
y1bar=apply(y1rs.mat,2,mean)
y2bar=apply(y2rs.mat,2,mean)

tstat=(y1bar-y2bar-mudiff)/sqrt(sy1sq/n1+sy2sq/n2)


alpha=1-conf.level # calculating alpha
#ci=quantile(xstat,c(alpha/2,1-alpha/2))# Nice way to form a confidence interval
pvalue=ifelse(test=="two",length(tstat[tstat>abs(tcalc) | tstat < -abs(tcalc)])/iter,
ifelse(test=="upper",length(tstat[tstat>tcalc])/iter,
length(ytstat[tstat<tcalc])/iter))

h=hist(tstat,plot=FALSE)
mid=h$mid
if(test=="two"){
ncoll=length(mid[mid<= -abs(tcalc)])
ncolr=length(mid[mid>=  abs(tcalc)])
col=c(rep("Green",ncoll),rep("Gray",length(mid)-ncoll-ncolr),rep("Green",ncolr))
}
hist(tstat,col=col,freq=FALSE)
#segments(ci[1],0,ci[2],0,lwd=2)

return(list(pvalue=pvalue))
#return(list(pvalue=pvalue,tcalc=tcalc,n=n,x=x,test=test,ci=ci))
}

boot2pval(x1=y,x2=x)
boot2pval(x1=y,x2=x,mudiff=2)

```

## Task 6

```{r}
set.seed(30); x2=rnorm(15,mean=10,sd=4)
set.seed(40); y2=rnorm(20,mean=12,sd=4)
boot2pval(x1=y2,x2=x2)

boot2pval(x1=y2,x2=x2,mudiff=2)
```

## Task 7

For line A, t.test with x1 and mu = 23 will perform a one sample t test for the mean of 23. This test finds the p-value and confidence interval for the mean of x1. If the p-value is less than 0.05 so H0 is rejected.

For line B, it seems like it is a comment describing what line a is doing. 

For line C, this seems like what the result of t.test is from line A. It is also similar to the outputs received previously. 

For line D, it is describing the alternative hypothesis where the true mean is not equal to 23. This is similar to H1 : u doesn't equal 23 as its the alternative hypothesis.

For line E, this is indicating that the data is being performed at a 95 percent confidence interval for the mean so if the mean is not in the interval from using a 95 percent confidence then the hypothesis could be rejected.

For line F, this looks like the interval received most likely from using that 95 percent confidence. Given earlier that mu = 23 it wont fit in this interval as it starts at 23.3.

For line G, this is showing that the sample mean is greater than the hypothesis mean from the beginning being 23. This one is 25.29 and exists within the interval. 

## Task 8

```{r}
library(FALL24DOUG0060)

FALL24DOUG0060::mypvalue(t0 = 2, xmax = 3, n = 10, alpha = 0.05)
```



