---
title: "R package"
author: "Andrew Douglass"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction



Congratulations for getting this far. To complete the evidence that you have made a working R package please answer the following by using the supplied rmd  and upload the knitted html to CANVAS.

I will assume your package is called **FALL24DOUG0060** and you have packaged and documented functions `myf1`, `myf2`, `myf3` and `myf4`.

Your actual package name and functions will obviously be called something else.

# First function

Please place your first function in the r chunk in the following way:

Once you are ready please remove the option `eval = FALSE`

```{r first}
library(FALL24DOUG0060)
FALL24DOUG0060::myclt(n = 100, iter = 1000, a = 0, b = 5)

```


# Second function

Please place your second function in the r chunk in the following way:

Once you are ready please remove the option `eval = FALSE`

```{r second, eval = FALSE}
set.seed(20)
x = rnorm(20, mean = 10, sd = 15)
FALL24DOUG0060::myci(x)

```


# Third function

Please place your third function in the r chunk in the following way:

Once you are ready please remove the option `eval = FALSE`

```{r third, eval = FALSE}
FALL24DOUG0060::myncurve(mu = 25, sigma = 5, a = 30)

```


# Fourth function

Please place your fourth function in the r chunk in the following way:

Once you are ready please remove the option `eval = FALSE`

```{r fourth, eval = FALSE}
FALL24DOUG0060::ntickets(n = 150, gamma = 0.05, p = 0.90)

```

# The github repository



This is the address of my GITHUB repository:
<https://github.com/MATHSTATSOU/Intro2R.git>

Replace this with your own address.
